%
% the CPPSOM Style Guide
%
% $Id: styleguide.tex 169 2008-29-10 17:41:48Z arne.bergmann $
%
\documentclass{scrartcl}
\usepackage{listings}
\begin{document}

\section{CPPSOM Style Guidelines}
\label{sec:cppsom-style-guid}

\subsection{Common Rules}
\label{sec:common-rules}
\begin{itemize}
\item A line shall have at most 80 columns.
\item Each file shall start with copyright and licensing information.
  Just before both, the string \texttt{\$Id\$} shall be placed as a
  comment.
\item The sole language of coding and documentation is
  \emph{English}.
\item A file shall end with a new line.
\end{itemize}


\subsection{Rules for C++}
\label{sec:rules-cpp}

The following rules apply to all CPP sourcecode files.

\begin{itemize}
\item The language used is C++ with STL. The target compiler is GNU
 g++ 4. The author strongly promotes maintaining code compatibility
 with Microsoft's Visual Studio 2008 cl compiler.
\item \emph{Avoid code duplication.}
\item All \lstinline|#endif|-statement shall be appended with a
  name. For mactros encpasulation a header file, this is the filename
  formatted as \lstinline|FILENAME_H_|.
\lstinline|#ifdef ... #endif| blocks shall start a new indentation
level.
\begin{lstlisting}
// do this
#ifdef DEBUG
    i = 10;
#endif DEBUG

// not this
#ifdef DEBUG
    i = 10;
#endif /*DEBUG*/

// nor this
#ifdef DEBUG
i = 10;
#endif
\end{lstlisting}

\item The following applies to all header file class definitions
\begin{itemize}
\item Every CPPSOM header shall be marked with \emph{\#pragma once}
 AND be marked with \emph{\#ifndef/\#define/\#endif} guards.
\item CPPSOM classes used in header files shall be defined using
 forward declaration. The \#include is to be done in the .cpp file.
 This does not apply to the base class of the class.
\begin{lstlisting}
//do this
#include <vmobjects/VMObject.h>

class VMArray;

class MyVMObject : VMObject
{
public:
    MyVMObject(VMArray*) : VMObject();
    /* ... more public members */
private:
		VMArray* embeddedArray;
}

//not this
#include <vmobjects/VMObject.h>

#include <vmobjects/VMArray>

class MyVMObject : VMObject
{
public:
    MyVMObject(VMArray*) : VMObject();
    /* ... more public members */
private:
		VMArray* embeddedArray;
}
\end{lstlisting}
\item There shall be exactly one space between the word class and
 the class' name.
\item There shall be exactly one space before and after the colon
 ':' that separates the class name from the name of the base class.
\item All class definitions should state members and methods in order
 of their visibility. Start with public, then protected and finally
 private. The visibility qualifiers shall be at indentation level 0.
\item Member declaration within one visibility should be done in the
 order: static methods, static fields), constructors, methods, fields.
\item Methods shall only be implemented in the header if it is a
 getter or setter function without further functionality besides
 getting or setting the value of a member variable.
\end{itemize}
\item There shall be exactly one space between a method's type and
  its name.
\begin{lstlisting}
// do this
void SomeClass::SomeMethod(String fmt, ...)
{
    ...
}

// not this
void    SomeClass::SomeMethod(String fmt, ...)
{
    ...
}
\end{lstlisting}
\item When definig methods, in a method signature, the opening
  brace \lstinline|{| shall be on the  next line after the corresponding
  closing paren \lstinline|)| at the same indentation level as the 
  method's return type.
\begin{lstlisting}
// do this
void SomeClass::SomeMethod(char *fmt, ...) 
{
    ...
}

// not this
void SomeClass::SomeMethod(String fmt, ...) {
    ...
}

// nor this
void SomeClass::SomeMethod(String fmt, ...){
    ...
}

// and not this
void SomeClass::SomeMethod(String fmt, ...)
        {
    ...
        }
\end{lstlisting}
\item An opening paren \lstinline|(| shall not be appended, a closing
  paren \lstinline|)| shall not be prepended with a space.
\begin{lstlisting}
// do this
bool right = ((a || b) && (c ^^ d)) || (e > f);

// not this
bool right = ( ( a || b ) && ( c ^^ d ) ) || ( e > f ) ;
\end{lstlisting}
\item Functions shall not start with empty lines. Code or comments
  shall start immediately.
\begin{lstlisting}
// do this
void print(char *fmt, ...) 
{
    int i=0;
    ...
}

// not this
void print(char *fmt, ...) 
{

    int i=0;
    ...
}
\end{lstlisting}
\item Each function definition is preceded and followed by two blank lines.
\begin{lstlisting}
// do this


void foo1(...) 
{
    ...
}


void foo2(...) 
{
    ...
}


//not this
void foo1(...) 
{
    ...
}
void foo2(...) 
{
    ...
}
\end{lstlisting}
\item \lstinline|**| and \lstinline|*| shall not be unnecessarily
  surrounded by spaces.
\begin{lstlisting}
// do this
array = *another;

// not this
array = * another;
\end{lstlisting}
\item When casting, the closing paren \lstinline|)| after the target type shall not
  be appended by a space.
\begin{lstlisting}
// do this
foo = (char*)bar;

// not this
foo = (char*) bar;
\end{lstlisting}
\item Redundant parens shall be avoided.
\begin{lstlisting}
// do this
return 0 == (a - b);

// not this
return (0 == (a - b));
\end{lstlisting}
\item In pointer definitions, the asterisk \lstinline|*| shall be
  aligned to the type not to the variable's name.
\begin{lstlisting}
// do this
int* foo;

// not this
int *foo;

// nor this
int * foo;
\end{lstlisting}
\item Comments like \lstinline|//end of somefile.c| shall be avoided.
\item A space shall be placed behind commas \lstinline|,| in parameter
  lists, base class lists et al.
\begin{lstlisting}
// do this
foo(a, b, c);

// not this
foo(a,b,c);

// nor this
foo(a ,b ,c);
foo(a , b , c);
\end{lstlisting}
\item Binary operators, shall be surrounded by spaces.
\begin{lstlisting}
// do this
bool right = !((++a | b) && (c ^^ d));

// not this
bool right=! ((++ a|b)&&(c^^d));
\end{lstlisting}
\item \lstinline|#pragma mark| code separators are preceded and followed by two blank lines.
\begin{lstlisting}
//do this
void foo(...) 
{
    ...
}


#pragma mark Helper functions


int helper(...) 
{
    ...
}

// not this
void foo(...) 
{
    ...
}
#pragma mark Helper functions
int helper(...) 
{
    ...
}
\end{lstlisting}
\item Backslashes \lstinline|\| shall be prepended by a space if
  encountered on line endings.
  \begin{lstlisting}
// do this
#define MY_LONG_MACRO_FORMAT \
    int i; \
    int j; \
    int k

// not this
#define MY_LONG_MACRO_FORMAT\
    int i;\
    int j;\
    int k
\end{lstlisting}
\item There shall not be a space between \lstinline|if/for/while/anyFunctionName|
  and the following opening paren \lstinline|(|.
  \begin{lstlisting}
// do this
if(a == b) 
{
    while(x > 2) 
    {
        z = foo(x);
    }
}

// not this
if (a == b) 
{
    while (x > 2) 
    {
        z = foo (x);
    }
}
\end{lstlisting}
\item \lstinline|#define| blocks shall not end with semi colon
  \lstinline|;| if the macro defined will be used with semi colons
  later on. This apllies especially to the \lstinline|_FORMAT|
  macros.
  \begin{lstlisting}
// do this
#define MY_LONG_MACRO_FORMAT \
    int i; \
    int j; \
    int k

// not this
#define MY_LONG_MACRO_FORMAT \
    int i; \
    int j; \
    int k;
\end{lstlisting}
\item Function comments shall be written in \verb|javadoc|-style,
  i.~e. at least
  \begin{lstlisting}
/**
 * brief description of what the function does
 */
\end{lstlisting}They are \emph{immediately} followed by the function header without any blank lines.
\item Each block of line comments shall start and end with a empty comment
  line. Such blocks are also preceded and followed by two blank lines.
\begin{lstlisting}
// do this


//
// some useful commentary
//


// not this
// some less useful commentary
\end{lstlisting}
\item There are no TABs (i.~e. no ASCII character 09). The indentation
  width shall be 4 spaces.
\begin{lstlisting}[showspaces]
// do this
void foo(void) 
{
    int i;
    if(a == ...) 
    {
        while(true) 
        {
            ...
        }
        ...
    }
}
\end{lstlisting}
\lstinputlisting[showtabs]{tab.c}
\item Lines which are too long to fit the 80 columns shall be split, such
  that it remains sensefully grouped. For assignation, this means to try
  to keep the right side of it together.
  \begin{lstlisting}
// do this
int foo =
    object->SomeFunctionName(first_parameter, second_parameter, third_parameter);

// not this
int foo =  object->SomeFunctionName(first_parameter, second_parameter,
    third_parameter);
  \end{lstlisting}
\item When using includes from different CPPSOM-directories, they shall
  be grouped and parted by empty lines.
  \begin{lstlisting}
// do this
#include <memory/gc.h>

#include <misc/String.h>

#include <vmobjects/VMFrame.h>
#include <vmobjects/VMBigInteger.h>
#include <vmobjects/VMMethod.h>

#include <compiler/Parser.h>

// not this
#include <compiler/Parser.h>
#include <misc/String.h>
#include <vmobjects/VMBigInteger.h>
#include <vmobjects/VMMethod.h>
#include <memory/gc.h>
#include <vmobjects/VMFrame.h>
\end{lstlisting}
\end{itemize}

\subsection{Rules fom SOM}
\label{sec:rules-fom-som}

\subsection{Rules for Make et al.}
\label{sec:rules-make-et}

\begin{itemize}
\item \emph{Do not use recursive makes.}
\end{itemize}

\end{document}